---
title: "Bryophytes BIEN"
output: html_notebook
---

```{r}
library(BIEN)
library(ape)
library(maps) 
library(sp) 
require(dplyr)
require(maptools)
require(raster)
require(dismo)
require(sp)
require(rgdal)
require(mapdata)
require(mapproj)
```


#Creates command: BIEN_occurrence_higher_plant_group
```{r}
BIEN_occurrence_higher_plant_group <- function (higher_plant_group, cultivated = FALSE, only.new.world = TRUE, 
                                                print.query = FALSE, observation.type = FALSE, all.taxonomy = FALSE, 
                                                native.status = FALSE, political.boundaries = FALSE, ...) 
{
  is_char(higher_plant_group)
  is_log(cultivated)
  is_log(only.new.world)
  is_log(print.query)
  is_log(observation.type)
  is_log(all.taxonomy)
  is_log(native.status)
  is_log(political.boundaries)
  if (!cultivated) {
    cultivated_query <- "AND (is_cultivated = 0 OR is_cultivated IS NULL)"
    cultivated_select <- ""
  }
  else {
    cultivated_query <- ""
    cultivated_select <- ",is_cultivated,is_cultivated_in_region"
  }
  if (only.new.world) {
    newworld_query <- "AND is_new_world = 1 "
    newworld_select <- ""
  }
  else {
    newworld_query <- ""
    newworld_select <- ",is_new_world"
  }
  if (!all.taxonomy) {
    taxon_select <- ""
  }
  else {
    taxon_select <- "verbatim_family,verbatim_scientific_name,family_matched,name_matched,name_matched_author,higher_plant_group,taxonomic_status,scrubbed_author,"
  }
  if (!native.status) {
    native_select <- ""
  }
  else {
    native_select <- "native_status,native_status_reason,native_status_sources,isintroduced,native_status_country,native_status_state_province,native_status_county_parish,"
  }
  if (!observation.type) {
    observation_select <- ""
  }
  else {
    observation_select <- ",observation_type"
  }
  if (!political.boundaries) {
    political_select <- ""
  }
  else {
    political_select <- "country,state_province,county,locality,"
  }
  query <- paste("SELECT higher_plant_group,", taxon_select, native_select, 
                 political_select, "scrubbed_species_binomial, latitude, longitude,date_collected,datasource,dataset,dataowner,custodial_institution_codes,collection_code", 
                 paste(cultivated_select, newworld_select, observation_select), 
                 "\n                 FROM view_full_occurrence_individual \n                 WHERE higher_plant_group in (", 
                 paste(shQuote(higher_plant_group, type = "sh"), collapse = ", "), 
                 ")", paste(cultivated_query, newworld_query), " AND higher_plant_group IS NOT NULL AND (is_geovalid = 1 OR is_geovalid IS NULL) ORDER BY scrubbed_species_binomial;")
  if (print.query) {
    query <- gsub(pattern = "\n", replacement = "", query)
    query <- gsub("(?<=[\\s])\\s*|^\\s+|\\s+$", "", query, 
                  perl = TRUE)
    print(query)
  }
  return(BIEN_sql(query, ...))
}
```


#Runs Command and stores as Data Set
```{r}
bryophyte_occurrence <- BIEN_occurrence_higher_plant_group("bryophytes")
```


#Creates Command BIEN_taxonomy_higher_plant_group to access species names of bryophytes in BIEN
```{r}
BIEN_taxonomy_higher_plant_group <- function (higher_plant_group, print.query = FALSE, ...) 
{
  is_char(higher_plant_group)
  is_log(print.query)
  sql_select <- paste("SELECT DISTINCT higher_plant_group, \"class\", superorder, \"order\", scrubbed_family,scrubbed_genus,scrubbed_species_binomial,scrubbed_author,scrubbed_taxonomic_status")
  sql_from <- paste(" FROM bien_taxonomy")
  sql_where <- paste(" WHERE higher_plant_group in (", paste(shQuote(higher_plant_group, 
                                                                     type = "sh"), collapse = ", "), ") AND scrubbed_species_binomial IS NOT NULL")
  sql_order_by <- paste(" ORDER BY higher_plant_group,scrubbed_family,scrubbed_genus,scrubbed_species_binomial,scrubbed_author ")
  query <- paste(sql_select, sql_from, sql_where, sql_order_by, 
                 ";")
  if (print.query) {
    query <- gsub(pattern = "\n", replacement = "", query)
    query <- gsub("(?<=[\\s])\\s*|^\\s+|\\s+$", "", query, 
                  perl = TRUE)
    print(query)
  }
  return(BIEN_sql(query, ...))
}
```


#Creates Bryophyte Database
```{r}
bryophytes <- BIEN_taxonomy_higher_plant_group("bryophytes", print.query = FALSE)
```


#Importing SpeciesPresence
```{r}
SpeciesPresence <- read.csv("~/Downloads/SpeciesPresence_by_100km_Cell.csv")
```


#Replacing "_" with " " in SpeciesPrescence
```{r}
SpeciesPresence$Species <- gsub("_", " ", SpeciesPresence$Species)
```


#Filter Dataset
```{r}
FilteredOccurrence=SpeciesPresence[SpeciesPresence$Species %in% unique(bryophyte_occurrence$scrubbed_species_binomial), ]
View(FilteredOccurrence)
```

#Tally
```{r}
tally <- tally(group_by(FilteredOccurrence, CellID))
colnames(tally)[2] <- "Richness"
```

#Import Raster
```{r}
BIENblank <-raster("~/Downloads/blank_100km_raster.tif")
```


#Create Data with all Columns
```{r}
Moss_Richness <- numeric(15038)
Moss_Richness[tally$CellID] <- tally$Richness
```


#Plot Data
```{r}
RichnessRaster <- setValues(BIENblank, Moss_Richness)
plot(RichnessRaster, axes = F)
```


#Take off gray
```{r}
RichnessVector<-Moss_Richness
RichnessVector[which(RichnessVector==0)]=NA
RichnessRasterNA <- setValues(BIENblank, RichnessVector)
plot(RichnessRasterNA, axes = F)
```


#Import Phylogeny
```{r}
bryphy <- read.csv("~/Documents/Courses/Research/BIEN/Bryophyte Phylogeny.csv")
```


#Mosses <- subset(bryphy, Group=="Mosses")
#Liverworts <- subset(bryphy, Group=="Liverworts")
#Hornworts <- subset(bryphy, Group=="Hornworts")

Rename Columns
```{r}
bryophytes$Family <- bryophytes$scrubbed_family
bryophytes$Genus <- bryophytes$scrubbed_genus
bryophytes$Species <- bryophytes$scrubbed_species_binomial
```

Data Manipulation
```{r}
newbryophyte <- bryophytes[,c(10,11,12)]
#data1<- newdata[,c(4,5,6)]
```


#Add Group to Bryophyte data by family
```{r}
newbryophyte <- merge(newbryophyte, bryphy, by = "Family", all.x = TRUE)
```

#Add Group to Occurrence data by species
```{r}
FilteredOccurrence <- merge(FilteredOccurrence, newbryophyte, by = "Species", all.x= TRUE)
```

#Subset Occurrence data by group
```{r}
mossoccurrence <- subset(FilteredOccurrence, FilteredOccurrence$Group=="Mosses")
hornwortoccurrence <- subset(FilteredOccurrence, FilteredOccurrence$Group=="Hornworts")
liverwortoccurrence <- subset(FilteredOccurrence, FilteredOccurrence$Group=="Liverworts")
hornandliveroccurrence <- subset(FilteredOccurrence, FilteredOccurrence$Group=="Hornworts"|FilteredOccurrence$Group=="Liverworts")
```


#Mapping Mosses
#Tally
```{r}
require(dplyr)
MossTally <- tally(group_by(mossoccurrence, CellID))
colnames(MossTally)[2] <- "Richness"
```


#Import Raster
#BIENblank <-raster("~/Downloads/blank_100km_raster.tif")

#Create Data with all Columns
```{r}
Mosses_Richness <- numeric(15038)
Mosses_Richness[MossTally$CellID] <- MossTally$Richness
```


#Plot Data
```{r}
MossRichnessRaster <- setValues(BIENblank, Mosses_Richness)
plot(MossRichnessRaster)
MossRichnessVector<-Mosses_Richness
MossRichnessVector[which(MossRichnessVector==0)]=NA
MossRichnessRasterNA <- setValues(BIENblank, MossRichnessVector)
plot(MossRichnessRasterNA, axes = F)
```



#Trial of Mapping Hornworts

#Tally
```{r}
require(dplyr)
HornwortTally <- tally(group_by(hornwortoccurrence, CellID))
colnames(HornwortTally)[2] <- "Richness"
```

#Import Raster
#BIENblank <-raster("~/Downloads/blank_100km_raster.tif")

#Create Data with all Columns
```{r}
Hornwort_Richness <- numeric(15038)
Hornwort_Richness[HornwortTally$CellID] <- HornwortTally$Richness
```


#Plot Data
```{r}
HornwortRichnessRaster <- setValues(BIENblank, Hornwort_Richness)
plot(HornwortRichnessRaster)
HornwortRichnessVector<-Hornwort_Richness
HornwortRichnessVector[which(HornwortRichnessVector==0)]=NA
HornwortRichnessRasterNA <- setValues(BIENblank, HornwortRichnessVector)
plot(HornwortRichnessRasterNA, axes = F)
```


#Mapping Liverworts
#Tally
```{r}
require(dplyr)
LiverwortTally <- tally(group_by(liverwortoccurrence, CellID))
colnames(LiverwortTally)[2] <- "Richness"
```


#Import Raster
#BIENblank <-raster("~/Downloads/blank_100km_raster.tif")

#Create Data with all Columns
```{r}
Liverwort_Richness <- numeric(15038)
Liverwort_Richness[LiverwortTally$CellID] <- LiverwortTally$Richness
```


#Plot Data
```{r}
LiverwortRichnessRaster <- setValues(BIENblank, Liverwort_Richness)
plot(LiverwortRichnessRaster)
LiverwortRichnessVector<-Liverwort_Richness
LiverwortRichnessVector[which(LiverwortRichnessVector==0)]=NA
LiverwortRichnessRasterNA <- setValues(BIENblank, LiverwortRichnessVector)
plot(LiverwortRichnessRasterNA, axes = F)
```



#Trial of Mapping Hornworts and Liverworts

#Tally
```{r}
require(dplyr)
HornandLiverTally <- tally(group_by(hornandliveroccurrence, CellID))
colnames(HornandLiverTally)[2] <- "Richness"
```

#Import Raster
#BIENblank <-raster("~/Downloads/blank_100km_raster.tif")

#Create Data with all Columns
```{r}
HornandLiver_Richness <- numeric(15038)
HornandLiver_Richness[HornandLiverTally$CellID] <- HornandLiverTally$Richness
```


#Plot Data
```{r}
HornandLiverRichnessRaster <- setValues(BIENblank, HornandLiver_Richness)
plot(HornandLiverRichnessRaster)
HornandLiverRichnessVector<-HornandLiver_Richness
HornandLiverRichnessVector[which(HornandLiverRichnessVector==0)]=NA
HornandLiverRichnessRasterNA <- setValues(BIENblank, HornandLiverRichnessVector)
plot(HornandLiverRichnessRasterNA, axes = F)
 ```

#FUNCTIONS
#Group
```{r}
group.map <- function(x, ...){
  k <- subset(FilteredOccurrence, FilteredOccurrence$Group== x)
  b <- tally(group_by(k, CellID))
  colnames(b)[2] <- "Richness"
  BIENblank <-raster("~/Downloads/blank_100km_raster.tif")
  f <- numeric(15038)
  f[b$CellID] <- b$Richness
  c <- setValues(BIENblank, f)
  plot(c)
  g<-f
  g[which(g==0)]=NA
  h <- setValues(BIENblank, g)
  plot(h, axes = F)
}
```

Try it out:
```{r}
group.map("Mosses")
group.map("Hornworts")
group.map("Liverworts")
```


#Species
```{r}
species.map <- function(x, ...){
  k <- subset(FilteredOccurrence, FilteredOccurrence$Species== x)
  b <- tally(group_by(k, CellID))
  colnames(b)[2] <- "Richness"
  BIENblank <-raster("~/Downloads/blank_100km_raster.tif")
  f <- numeric(15038)
  f[b$CellID] <- b$Richness
  c <- setValues(BIENblank, f)
  plot(c)
  g<-f
  g[which(g==0)]=NA
  h <- setValues(BIENblank, g)
  plot(h)
}
```

Try it out:
```{r}
species.map("Anacolia laevisphaera")
species.map("Splachnum sphaericum")
species.map("Physcomitrella patens")
```




#Family Function
```{r}
family.map <- function(x, ...){
  k <- subset(FilteredOccurrence, FilteredOccurrence$Family== x)
  b <- tally(group_by(k, CellID))
  colnames(b)[2] <- "Richness"
  BIENblank <-raster("~/Downloads/blank_100km_raster.tif")
  f <- numeric(15038)
  f[b$CellID] <- b$Richness
  c <- setValues(BIENblank, f)
  plot(c)
  g<-f
  g[which(g==0)]=NA
  h <- setValues(BIENblank, g)
  plot(h)
}
```

Try it out:
```{r}
family.map("Sphagnaceae")
family.map("Anthocerotaceae")
```


#Create Dataset showing Range
```{r}
Range <- tally(group_by(FilteredOccurrence, Species))
Filter <- merge(Range, FilteredOccurrence, by = "Species")
```


#Find Median of CellIDs
```{r}
filter.2 <- subset(Filter, select = c("n", "CellID")) %>%
     group_by(CellID) %>%
     summarize(Avg = median(n))
```


#Create Range Map
```{r}
BIENblank <-raster("~/Downloads/blank_100km_raster.tif")
Moss_Range.3 <- numeric(15038)
Moss_Range.3[filter.2$CellID] <- filter.2$Avg
RangeRaster <- setValues(BIENblank, Moss_Range.3)
plot(RangeRaster)
```

#Take off gray
```{r}
RangeVector<-Moss_Range.3
RangeVector[which(RangeVector==0)]=NA
RangeRasterNA <- setValues(BIENblank, RangeVector)
plot(RangeRasterNA)
```



